cvx-flats

### 2020-04-22 neurips-20/cvx-flats.tex
Fixed Thm 1 and added some general writing

@@

\subsubsection{Elicitation Complexity}
When properties are vector-valued (i.e.$\, \Gamma: \simplex \to \reals^d$),~\cite{frongillo2015elicitation} introduce the notion of \emph{elicitation complexity}.

@@

\jessie{Note from 2018 neurips paper that 1-elicitable iff 1-convex elicitable for continuous properties and how this differs.}

\subsection{Consistency and calibration for convex losses}\label{subsec:convex-surrogates}
\cite{zhang2004statistical,lin2004note,bartlett2006convexity,tewari2007consistency} form a characterization of consistent, calibrated surrogates for classification problems.
In particular,~\cite{bartlett2006convexity} show that if a convex surrogate is differentiable, minimal at $0$, and nonnegative, there is a consistent surrogate for binary classification problems, and ~\cite{tewari2007consistency} generalizes this result for multiclass classification. \jessie{add \cite{ramaswamy2016convex, ramaswamy2018consistent} notes}

\cite{steinwart2007compare} generalizes the study of consistent losses from the classification setting \ldots \jessie{Need to come back to this.  area exam Jessie looked at this briefly, but current Jessie is a whole different person.}

@@

\jessiet{Excess risk definition on its own}

@@

Proposition~\ref{prop:consistent-implies-indir-elic} allows us to draw the connection between property elicitation and consistent losses.
Moreover, combining Proposition~\ref{prop:consistent-implies-indir-elic} and Theorem~\ref{thm:cvx-flats}, we provide new conditions for finding lower bounds on \emph{convex calibration dimension} for finite outcome settings in Section~\ref{sec:finite-calib}.

\begin{proposition}\label{prop:consistent-implies-indir-elic}
	If a surrogate and link pair $(L, \psi)$ is consistent with respect to $\ell$, then $L$ indirectly elicits $\gamma := \prop{\ell}$.
\end{proposition}
\begin{proof}
Since consistency must hold for all $D$ over $\X \times \Y$, let us focus on distributions $D$ such that the probability of any $x \in \X$ is a point mass, allowing us to focus on conditional distributions $p \in \simplex$ such that $p = Pr[Y = y | X = x]$, allowing us to be agnostic to $\X$.
Now, instead of focusing on sequences of optimal hypotheses, we can focus on optimal reports when given $x \in \X$ if the hypothesis class $\F$ is sufficiently rich (i.e. the optimal hypothesis is in the hypothesis class for all $x \in \X$.)
 \jessie{FINISH}
\end{proof}

\begin{remark}
	Either definition of consistency ties to indirect elicitation
\end{remark}

@@

\begin{definition}[Flat]
	A \emph{flat} $F$ of $\reals^n$ is an affine subspace of $\reals^n$.
  \raft{Would suggest defining a bit more formally after, like ``In other words, $F\subseteq \reals^d$ can be written $F=\{x\in\reals^n : Wx + b = 0\}$ for some $W\in\reals^{d\times n}$ and $b\in\reals^d$ where $d\leq n$.  Then you can define the dimension of $F$ easily as $d$, assuming the matrix is full rank}
\end{definition}

@@

\begin{theorem}\label{thm:cvx-flats}
	Suppose we are given a property $\gamma$.
  \raft{Made logic more direct}
	If, for any $p \in \simplex$, and $r\in\gamma(p)$,
	there is no $(n - d - 1)$-dimensional flat $F$ satisfying both $F \cap \simplex \subseteq \gamma_r$ and $p \in F$, then there is no convex surrogate loss $L : \reals^d \times \Y \to \reals$ that indirectly elicits $\gamma$.
\end{theorem}
\begin{proof}[Proof intuition]
	We prove the contrapositive: if there is a surrogate loss $L$ indirectly eliciting $\gamma$, then for all $p \in \simplex$ and $r \in \R$ such that $r \in \gamma(p)$, there is an $(n-d-1)$-dimensional flat $F$ such that $F \cap \simplex \subseteq \gamma_r$ and $p \in F$.
  \raft{Nice way to start.}

	Suppose we have $L:\reals^d \times \Y \to \reals$ eliciting $\Gamma$ and link $\psi : \reals^d \to \R$ such that for all $u \in \reals^d$, we have $\Gamma_u \subseteq \gamma_{\psi(u)}$, i.e., $L$ indirectly elicits $\gamma$.
	Fix $p \in \simplex$, and take any $u \in \Gamma(p)$ such that $\psi(u) = r \in \gamma(p)$.
	We know that $u \in \Gamma(p) \iff \vec 0 \in \partial L(u; p) \iff \vec 0 \in \oplus_y p_y \partial L(u,y)$, where $\oplus$ denotes the nonnegative weighted Minkowski sum by convex analysis results \jessie{And mild assumptions on the loss?}\raft{It's a good exercise to look up the necessary results in e.g. Rockafellar (let me know if you don't have a copy; I can send a PDF).  Note that $L(\cdot,y)$ is defined on all of $\reals^d$ for all $y$, which is key.  I'd recommend spelling this out in 2-3 sentences.}.

	Observe that the Minkowski sum $\oplus_y p_y \partial L(u,y) = \{p_1 x_1 + \ldots + p_n x_n : x_1 \in \partial L(u,1), \ldots, x_n \in \partial L(u,n) \}$.
	In order for $\vec 0$ to be in this $p$-weighted Minkowski sum, we then have that each choice of $x_i$ is a flat for $\partial L(u, i)$.
	Moreover, with $p$ fixed, we know there is a vector of witnesses $W = [x_1, \ldots, x_n]$ such that $\sum_y p_y x_y = \vec 0$.
  \raft{State the dimensions of $W$}

	\jessiet{This next paragraph is mostly me spelling it out for myself}
  \raft{Don't understand; this is necessary to complete the proof!  (So good thing you have it.)}
	Observe that $F := \ker(W)$ is at most $(n-d-1)$ dimensions in probability space and contains $p$ by construction.
  \raft{Justify the dimension.  Hint: $F$ contains $\vec 0$.  You'll probably want to define some $F'$ by adding a certain something to $W$ -- check out the FSD paper.}
	We also claim that $F \cap \simplex \subseteq \gamma_r$.
	To see this, consider $q \in F \cap \simplex$, but $q \not \in \gamma_r$.
	That would mean that $Wq = \vec 0$, which implies $\vec 0 \in \partial L(u;q)$, implying $q \in \Gamma_u$.
	Therefore, $q \not \in \gamma_r$ would contradict $L$ indirectly eliciting $\gamma$.

		%%COMMENTED 04.16.20
%	Observe that the weighted Minkowski sum is in $d$-dimensional space as $L$ is defined on $\reals^d$.
%	As $\partial L(u;p)$ is defined in $n$-dimensional space over $\simplex$, the flats of $L(u;p)$ are at least $(n-d-1)$-dimensional as the subgradient is in $d$-dimensional space, and the affineness of being a flat allows us to remove one additional dimension.
%	Let $F$ be an $(n-d-1)$-dimensional flat to $L(u;p)$.
%	For all $q \in F$, we must have $u \in \Gamma(q)$ since \jessie{JUSTIFICATION} there is a bijection from $F \in \partial L(u;p)$ to $\Gamma(p)$. Therefore, $q \in F \implies u \in \Gamma(q)$ as well.
%	Therefore, we have $F \cap \simplex \subseteq \Gamma_u \subseteq \gamma_r$, where the latter follows from indirect elicitation.
%%end Commented 4.16.20

%	We claim there exists a matrix $W \in \reals^{d+1 \times n}$ defining the flat $F = \{q \in \reals^n : Wq = \vec 0 \}$; $F$ is then $(n-(d+1))$-dimensional by the rank-nullity theorem, so we just need to show such a $W$ also keeps $F \cap \simplex \subseteq \gamma_r$.
%	As $\Gamma_u \subseteq \gamma_r$, it is sufficient to show $F \cap \simplex \subset \Gamma_u$.
%	 
%	Construct $W$ by taking each column to describe $\partial L(u, y)$ for each $y \in \Y$.
%	Since $\partial L(u, y)$ is $d$-dimensional, the affine subspace containing each subgradient can be described with $d+1$ constraints. \jessie{??}
%	We can see that $p \in F$ since $\vec 0 \in \oplus_y p_y \partial L(u, y) \implies Wp = \vec 0$ by construction of $W$.
%	Moreover, if there was a $\hat q \in \simplex \setminus \gamma_r$ such that $W \hat q = \vec 0$, then we would contradict the indirect elicitation of $\gamma$ since we would not have $\Gamma_{u} \subseteq \gamma_{\psi(u)}$ since we would have $\hat q \in \Gamma_u$ but $\hat q \not \in \gamma_{\psi(u)}$.
%	Therefore,  we have an $(n - d - 1)$-dimensional flat $F$ such that $F \cap \simplex \subseteq \gamma_r$ and $p \in F$ for all $p \in \simplex$ and $r \in \gamma(p)$.
\end{proof}

>>>>

\cite{finocchiaro2018convex} are among the first to consider a characterization of \emph{convex} elicitable properties, in which they find that all continuous, nowhere locally constant elicitable properties (in finite outcome settings) are elicitable by a \emph{convex} loss.
However, their assumptions are more restrictive than ours; they only consider losses defined on the real line (see Section~\ref{subsec:elic-cplx} below for our generalization) and assume the properties to be identifiable: an assumption we do not need here.

@@

\subsubsection{Elicitation Complexity}\label{subsec:elic-cplx}
When properties are vector-valued (i.e.$\, \Gamma: \simplex \to \reals^d$),~\cite{frongillo2015elicitation} introduce the notion of \emph{elicitation complexity}.

@@

\subsection{Consistency and calibration for convex losses}\label{subsec:convex-surrogates}
\cite{zhang2004statistical,lin2004note,bartlett2006convexity,tewari2007consistency} form a characterization of consistent, calibrated surrogates for classification problems.
In particular,~\cite{bartlett2006convexity} show that if a convex surrogate is differentiable, minimal at $0$, and nonnegative, there is a consistent surrogate for binary classification problems, and ~\cite{tewari2007consistency} generalizes this result for multiclass classification. 
\cite{ramaswamy2016convex} further shows necessary and sufficient conditions for finite prediction problems to be consistent that can be applied to general loss matrices (i.e., problems where both the report and outcome set are finite.)
They additionally introduce a notion of \emph{convex calibration} dimension similar to elicitation complexity mentioned above; we discuss this further in Section~\ref{sec:finite-calib}.
Their calibration results then yield consistent surrogates for finite prediction problems such as hierarchical classification from~\cite{ramaswamy2015hierarchical} and classification with an abstain option studied by~\cite{ramaswamy2018consistent}.

\cite{steinwart2007compare} generalizes the study of consistent losses from the classification setting \ldots \jessiet{Need to come back to this.  area exam Jessie looked at this briefly, but current Jessie is a whole different person.}

@@

\jessie{Excess risk definition on its own?  Followed by statement that satisfying excess risk bound implies consistency.}

@@

Proposition~\ref{prop:consistent-implies-indir-elic} below allows us to understand the connection between property elicitation and consistent losses.
Moreover, combining Proposition~\ref{prop:consistent-implies-indir-elic} and Theorem~\ref{thm:cvx-flats}, we provide new conditions for finding lower bounds on \emph{convex calibration dimension} for finite outcome settings in Section~\ref{sec:finite-calib}.
In infinite outcome settings, we can generalize Theorem~\ref{thm:cvx-flats}\jessie{hopefully...} in order to answer an open question about convex elicitation complexity posed by~\cite{frongillo2015elicitation}.

\begin{proposition}\label{prop:consistent-implies-indir-elic}
	If a surrogate and link pair $(L, \psi)$ is consistent with respect to $\ell$, then $L$ indirectly elicits $\gamma := \prop{\ell}$.
\end{proposition}
\begin{proof}
Since consistency must hold for all $D$ over $\X \times \Y$, let us focus on distributions $D$ such that the probability of any $x \in \X$ is a point mass, allowing us to focus on conditional distributions $p \in \simplex$ such that $p = Pr[Y = y | X = x]$, allowing us to be agnostic to $\X$.
Now, instead of focusing on sequences of optimal hypotheses, we can focus on optimal reports when given $x \in \X$ if the hypothesis class $\F$ is sufficiently rich (i.e. the optimal hypothesis is in the hypothesis class for all $x \in \X$.)

If the infimum of $L$ is always attained in report space, then for

 \jessie{FINISH}
\end{proof}

\begin{remark}
	Both definitions of consistency tie to indirect elicitation
\end{remark}

@@

\begin{definition}[Flat]
	A \emph{flat} $F$ of $\reals^n$ is an affine subspace of $\reals^n$.
	In other words, $F\subseteq \reals^n$ can be written $F=\{x\in\reals^n : Wx + b = 0\}$ for some $W\in\reals^{d\times n}$ and $b\in\reals^d$ where $d\leq n$.
	By the rank-nullity theorem, the dimension of the flat $F$ is the nullity of $W$ by construction, and is given by $f := n - \mathrm{rank}(W)$.
	If $W$ is full-rank, then $f = n - d$.
%  \raft{Would suggest defining a bit more formally after, like ``In other words, $F\subseteq \reals^d$ can be written $F=\{x\in\reals^n : Wx + b = \vec 0\}$ for some $W\in\reals^{d\times n}$ and $b\in\reals^d$ where $d\leq n$.  Then you can define the dimension of $F$ easily as $d$, assuming the matrix is full rank}
\end{definition}
Unless otherwise mentioned, we will assume that $W$ is full rank.

@@

\begin{theorem}\label{thm:cvx-flats}
	Suppose we are given a property $\gamma$ and distribution $p \in \simplex$.
  \raft{Made logic more direct \jessie{Done?}}
	For all $r\in\gamma(p)$, if there is no $(n - d-1)$-dimensional flat $F$ containing $p$ so that $F \cap \simplex \subseteq \gamma_r$, then there is no convex surrogate loss $L : \reals^d \times \Y \to \reals$ that indirectly elicits $\gamma$.
\end{theorem}
\begin{proof}[Proof intuition]
	We prove the contrapositive: if there is a surrogate loss $L$ indirectly eliciting $\gamma$, then for all $p \in \simplex$ and $r \in \R$ such that $r \in \gamma(p)$, there is an $(n-d-1)$-dimensional flat $F$ such that $F \cap \simplex \subseteq \gamma_r$ and $p \in F$.
%  \raft{Nice way to start.}

	Suppose we have $L:\reals^d \times \Y \to \reals$ eliciting $\Gamma$ and link $\psi : \reals^d \to \R$ such that for all $u \in \reals^d$, we have $\Gamma_u \subseteq \gamma_{\psi(u)}$, i.e., $L$ indirectly elicits $\gamma$.
	Fix $p \in \simplex$, and take any $u \in \Gamma(p)$ such that $\psi(u) = r \in \gamma(p)$.
	We know that $u \in \Gamma(p) \iff \vec 0 \in \partial L(u; p) \iff \vec 0 \in \oplus_y p_y \partial L(u,y)$, where $\oplus$ denotes the nonnegative weighted Minkowski sum~\cite[Theorem 4.1.1]{hiriart2012fundamentals}.  (Observe that in order for $L$ to be convex, $L(\cdot, y)$ must be convex for all $y \in \Y$.) \raft{It's a good exercise to look up the necessary results in e.g. Rockafellar (let me know if you don't have a copy; I can send a PDF).  Note that $L(\cdot,y)$ is defined on all of $\reals^d$ for all $y$, which is key.  I'd recommend spelling this out in 2-3 sentences.}.

	Observe that the Minkowski sum $\oplus_y p_y \partial L(u,y) = \{p_1 x_1 + \ldots + p_n x_n : x_1 \in \partial L(u,1), \ldots, x_n \in \partial L(u,n) \}$.
	In order for $\vec 0$ to be in this $p$-weighted Minkowski sum, we then have that each choice of $x_i$ is a flat for $\partial L(u, i)$.
	Moreover, with $p$ fixed, we know there is a vector of witnesses $W' = [x_1, \ldots, x_n] \in \reals^{d\times n}$ such that $\sum_y p_y x_y = \vec 0$.
%  \raft{State the dimensions of $W$}

%  \raft{Don't understand; this is necessary to complete the proof!  (So good thing you have it.)}
	Observe that $F' := \ker(W')$ is at most $(n-d)$ dimensions in $\reals^n$ and contains $p$ by construction.
	Since we are only concerned with a flat intersected with the simplex being a subset of $\gamma_r$, we can reduce the dimension by adding an affine simplex constraint. 
	That is, take $F := \{x \in \reals^n : W x + b = \vec 0\}$, where we have $W := [W'; \mathbbm{1}^T] \in \reals^{(d+1) \times n}$ and $b \in \reals^{d+1}$.
	The dimension of $F$ is then $n - \mathrm{rank}(W)$, which is $n - (d+1)$ if $W$ is full rank.
  \raft{Justify the dimension.  Hint: $F$ contains $\vec 0$.  You'll probably want to define some $F'$ by adding a certain something to $W$ -- check out the FSD paper.}
	We also claim that $F \cap \simplex \subseteq \gamma_r$.
	To see this, consider $q \in F$, but $q \not \in \gamma_r$.
	That would mean that $Wq + b = \vec 0 \iff W'q + b' = \vec 0$ (since $q \in \simplex$ satisfies the simplex constraints).
	This in turn implies $\vec 0 \in \partial L(u;q)$, implying $q \in \Gamma_u$.
	Therefore, $q \not \in \gamma_r$ would contradict $L$ indirectly eliciting $\gamma$.
\end{proof}
